Disparate views of disparity

### 2017-12-06 Analysis/disparity_methods.Rmd
data cleaning pipeline

I collected the data from Google scholar per year for the key words "morphological disparity".
I analysed all 100 <!--@@@ TG: check if I stick to this value or not --> papers for each year and collected the following data:

 * The paper details
 * Whether or not the paper contained a disparity analysis (for consistency)
 * The focal biological group
 * The type of data (3D landmarks, discrete data, etc...)
 * Whether the full organism was used or not.
 * How was the morphospace implicitly defined?
 * How was disparity implicitly defined?
 * Which test was applied for testing hypothesis?
 * Was phylogeny taken into account?

@@

## Sorting the data to categorise it

The categories for each variable are:

```{r}
## Categories definition
data_type <- c("3D landmarks", "2D landmarks", "continuous", "discrete", "body", "metabolic")
full_organism <- c("yes", "no")
morphospace <- c("PCA", "MDS", "PCO", "PCOA", "Raup")
disparity <- c("sum of variance", "sum of range", "product of variance", "product of range", "distance", "observed") #TG: Area from observed to estimated (DTT)
test <- c("anova", "model fitting", "canonical")
phylo <- c("yes", "no", "mapped")

@@

>>>>

I collected the data from the 500 first Google Scholar results for the key words "morphological disparity" per order of appearance (accessed 1st of Noveember 2017).
This resulted in 232 relevant papers for which I collected:

 * The focal biological group
 * The type of data (3D landmarks, discrete data, etc...)
 * Whether the full organism was used or not.
 * How was the morphospace explicitly defined?
 * How was disparity explicitly defined?
 * Which test was applied for testing hypothesis?
 * Was phylogeny taken into account?

In the other 268 papers disparity was mentioned but not measured.
This happened for papers falling in the following categories:

 * Out of topic
 * Morphological descriptions
 * Review paper
 * Paper not accessible (broken link or paywall)
 * Citation only

Additionally I collected the number of matches per year from 1960 to 2016 for the terms "morphological disparity" both in the text (fuzzy matching) or in the title (exact matching).

@@

# Occurrences of morphological disparity papers

This first analysis just shows that disparity analysis in morphology are steadily and rapidly increasing since the 2000s:

```{r, fig.width = 8, fig.height = 8}
disparity_fuzzy <- search_results[-c(1:2, 60:65),3]
disparity_exact <- search_results[-c(1:2, 60:65),2]

## Number of papers
# pdf("../Paper/Figures/GoogleScholarOccurences.pdf", width = 8, height = 8)
op <- par(bty = "n")
plot(seq(from = 1960, to = 2016), rev(disparity_fuzzy)/search_results[1,3], type = "l", xlab = "", ylab = "Relative number of papers per year", ylim = c(0, 0.1))
lines(seq(from = 1960, to = 2016), rev(disparity_exact)/search_results[1,2], type = "l", col = "grey")
par(op)
# dev.off()
```

Not much to say here but that it is a trendy topic!

# Disparate views of disparity

Disparity methods comes in all flavours!

## Sorting the variables

The variables can be sorted as follow:

```{r}

## Categories definition
data_type <- c("3D landmarks", "2D landmarks", "continuous", "discrete", "body", "metabolic")
full_organism <- c("yes", "no")
morphospace <- c("PCA", "MDS", "PCO", "PCOA", "Raup")
disparity <- c("sum of variance", "sum of range", "product of variance", "product of range", "distance", "observed") #TG: Area from observed to estimated (DTT)
test <- c("anova", "model fitting", "canonical", "confidence interval")
phylo <- c("yes", "no", "mapped")

@@

# Occurrences of morphological disparity papers

```{r}
disparity_fuzzy <- search_results[-c(1:2, 60:65),3]
disparity_exact <- search_results[-c(1:2, 60:65),2]

## Number of papers
pdf("../Writing/Figures/GoogleScholarOccurences.pdf", width = 8, height = 8)
op <- par(bty = "n")
plot(seq(from = 1960, to = 2016), rev(disparity_fuzzy)/search_results[1,3], type = "l", xlab = "", ylab = "Number of papers per year", ylim = c(0, 0.1), yaxt = "n")
lines(seq(from = 1960, to = 2016), rev(disparity_exact)/search_results[1,2], type = "l", col = "grey")
## Adding the axis
axis(2, at = seq(from = 0, to = 0.08, by = 0.02), labels = round(seq(from = min(disparity_fuzzy), to = max(disparity_fuzzy), length.out = 5)) )
axis(4, at = seq(from = 0, to = 0.1, by = 0.02), labels = round(seq(from = min(disparity_exact), to = max(disparity_exact), length.out = 6)), col = "grey", las = 2)

par(op)
dev.off()
```
